### 2020-09-21 submission/manuscript.tex
Edits for revision

\hypertarget{the-initial-gut-microbiota-and-response-to-antibiotic-perturbation-influence-clostridioides-difficile-colonization-in-mice}{%
\section{\texorpdfstring{The initial gut microbiota and response to
antibiotic perturbation influence \emph{Clostridioides difficile}
colonization in
mice}{The initial gut microbiota and response to antibiotic perturbation influence Clostridioides difficile colonization in mice}}\label{the-initial-gut-microbiota-and-response-to-antibiotic-perturbation-influence-clostridioides-difficile-colonization-in-mice}}

@@

We sequenced the 16S rRNA gene from fecal samples collected from these
mice after they acclimated to the University of Michigan animal housing
environment. We first examined the alpha diversity across the 6 sources
of mice. There was a significant difference in the richness (i.e.~number
of observed operational taxonomic units (OTUs)), but not Shannon
diversity index across the sources of mice (\emph{P}\textsubscript{FDR}
= 0.03 and \emph{P}\textsubscript{FDR} = 0.052, respectively; Fig. 1A-B
and Tables S1-2). Next, we compared the community structure of mice
(Fig. 1C). The source of mice and the interactions between the source
and cage effects explained most of the observed variation between fecal
communities (PERMANOVA combined R\textsuperscript{2} = 0.90, \emph{P}
\textless{} 0.001; Fig. 1C and Table S3). Mice that are co-housed tend
to have similar gut microbiotas due to coprophagy (41). Since mice
within the same source were housed together, it was not surprising that
the cage effect also contributed to the observed community variation.
There were some differences between the 2 experiments we conducted, as
the experiment and cage effects significantly explained the observed
community variation for the Schloss and Young lab mouse colonies (Fig.
S2A-B and Table S4). However, most of the vendors also clustered by
experiment (Fig. S2C-D, F), suggesting there was some community
variation between the 2 experiments within each source, particularly for
Schloss, Young, and Envigo mice (Fig. S2G-H). After finding differences
at the community level, we next identified the bacteria that varied
between sources of mice. There were 268 OTUs with relative abundances
that were significantly different between the sources (Fig. 1D and Table
S5). Though we saw differences between experiments at the community
level, there were no OTUs that were significantly different between
experiments within Schloss, Young, and Envigo mice (all \emph{P}
\textgreater{} 0.05). By using mice from six sources we were able to
increase the variation in the starting communities to evaluate in a
clindamycin-based CDI model.

\textbf{Clindamycin treatment renders all mice susceptible to \emph{C.
difficile} 630 colonization, but clearance time varies across sources.}
Clindamycin is frequently implicated with human CDIs (42) and was part
of the antibiotic treatment for the frequently cited 2008 CDI mouse
model (14). We have previously demonstrated mice are rendered
susceptible to \emph{C. difficile}, but clear the pathogen within 9 days
when treated with clindamycin alone (21, 43). All mice were treated with
10 mg/kg clindamycin via intraperitoneal injection and one day later
challenged with 10\textsuperscript{3} \emph{C. difficile} 630 spores
(Fig. 2A). The day after infection, \emph{C. difficile} was detectable
in all mice at a similar level (median CFU range: 2.2e+07-1.3e+08;
\emph{P}\textsubscript{FDR} = 0.15), indicating clindamycin rendered all
mice susceptible regardless of source (Fig. 2B). However, between 3 and
7 days post-infection, we observed variation in \emph{C. difficile}
levels across sources of mice (all \emph{P}\textsubscript{FDR} \(\le\)
0.019; Fig. 2B and Table S6). This suggested the source of mice was
associated with \emph{C. difficile} clearance. While the colonization
dynamics were similar between the two experiments, the Schloss mice took
longer to clear in the first experiment compared to the second and the
Envigo mice took longer to clear in the second experiment compared to
the first (Fig. S2A-B). The change in the mice's weight significantly
varied across sources of mice with the most weight loss occurring two
days post-infection (Fig. 2C and Table S7). There was also one Jackson
and one Envigo mouse that died between 1- and 3-days post-infection
during the second experiment. Mice obtained from Jackson, Taconic, and
Envigo tended to lose more weight, have higher \emph{C. difficile} CFU
levels and take longer to clear the infection compared to the other
sources of mice (although there was variation between experiments with
Schloss and Envigo mice). This was particularly evident 7 days
post-infection (Fig. 2B-C, Fig. S2C-D), when 57\% of the mice were still
colonized with \emph{C. difficile} (Fig. S2E). By 9 days post-infection
the majority of the mice from all sources had cleared \emph{C.
difficile} with the exception of 1 Taconic mouse from the first
experiment and 2 Envigo mice from the second experiment (Fig. 2B). Thus,
clindamycin rendered all mice susceptible to \emph{C. difficile} 630
colonization, regardless of source, but there was significant variation
in disease phenotype across the sources of mice.

\textbf{Clindamycin treatment alters bacteria in all sources, but a
subset of bacterial differences across sources persists.} Given the
variation in fecal communities that we observed across breeding
colonies, we hypothesized that variation in \emph{C. difficile}
clearance would be explained by community variation across the 6 sources
of mice. As expected, clindamycin treatment decreased the richness and
Shannon diversity across all sources of mice (Fig. 3A-B). Interestingly,
significant differences in diversity metrics between sources emerged
after clindamycin treatment, with Charles River mice having higher
richness and Shannon diversity than most of the other sources
(\emph{P}\textsubscript{FDR} \textless{} 0.05; Fig 3A-B and Tables
S1-2). The clindamycin treatment decreased the variation in community
structures between sources of mice. The source of mice and the
interactions between source and cage effects explained almost all of the
observed variation between communities (combined R\textsuperscript{2} =
0.99, \emph{P} \textless{} 0.001; Fig. 3C and Table S3). However, there
were only 18 OTUs with relative abundances that significantly varied
between sources (Fig. 3D and Table S8). Next, we identified the bacteria
that shifted after clindamycin treatment, regardless of source by
analyzing paired fecal samples from mice that were collected at baseline
and after clindamycin treatment. We identified 153 OTUs that were
altered after clindamycin treatment in most mice (Fig. 3E and Table S9).
When we compared the list of significant clindamycin impacted bacteria
with the bacteria that varied between sources post-clindamycin, we found
4 OTUs that were shared between the lists (\emph{Enterobacteriaceae (OTU
1)}, \emph{Lachnospiraceae} (OTU 130), \emph{Lactobacillus} (OTU 6),
\emph{Enterococcus} (OTU 23); Fig. 3D-E and Tables S8-9). Importantly,
some of the OTUs that varied between sources also shifted with
clindamycin treatment. For example, \emph{Proteus} increased after
clindamycin treatment (Fig. 3D), but only in Taconic mice.
\emph{Enterococcus} was primarily found only in mice purchased from
commercial vendors and also increased in relative abundance after
clindamycin treatment (Fig. 3D). These findings demonstrate that
clindamycin had a consistent impact on the fecal bacterial communities
of mice from all sources and only a subset of the OTUs continued to vary
between sources.

@@

\textbf{Baseline, post-clindamycin, and post-infection community data
can predict mice that will clear \emph{C. difficile} by 7 days
post-infection.} After identifying taxa that varied between sources,
changed after clindamycin treatment, or both, we determined which taxa
were influencing the variation in \emph{C. difficile} colonization at
day 7 (Fig. 2B, Fig. S2C). We trained three L2-regularized logistic
regression models with either input bacterial community data from the
baseline (day = -1), post-clindamycin (day = 0), or post-infection (day
= 1) timepoints of the experiment to predict \emph{C. difficile}
colonization status on day 7 (Fig. S3A-B). All models were better at
predicting \emph{C. difficile} colonization status on day 7 than random
chance (all \emph{P} \textless{} 0.001, Table S12). The model based on
the post-clindamycin (AUROC = 0.78) community OTU data performed
significantly better than the baseline (AUROC = 0.72) or the
post-infection (AUROC = 0.67) models (\emph{P}\textsubscript{FDR}
\textless{} 0.001 for pairwise comparisons; Fig. S3C and Table S13).
Thus, we were able to use bacterial relative abundance data from the
time of \emph{C. difficile} challenge to differentiate mice that had
cleared \emph{C. difficile} before day 7 from the mice still colonized
with \emph{C. difficile} at that timepoint. This result suggests that
the bacterial community's response to clindamycin treatment had the
greatest influence on subsequent \emph{C. difficile} colonization
dynamics.

To examine the bacteria that were driving each model's performance, we
selected the 20 OTUs that had the highest absolute feature weights in
each of the 3 models (Table S14). First, we looked at OTUs from the
model with the best performance, which was based on the post-clindamycin
treatment (day 0) bacterial community data. Out of the 10 highest ranked
OTUs, 7 OTUs were associated with \emph{C. difficile} colonization 7
days post-infection (\emph{Bacteroides}, \emph{Escherichia/Shigella}, 2
\emph{Lachnospiraceae}, \emph{Lactobacillus}, \emph{Porphyromonadaceae},
and \emph{Ruminococcaceae}), while 3 OTUs were associated with clearance
(\emph{Enterobacteriaceae}, \emph{Lachnospiraceae},
\emph{Porphyromonadaceae}; Fig. 5A). Next, we examined whether any of
the top 20 ranked OTUs from the post-clindamycin (day 0) model were also
important in the other 2 classification models based on baseline (day
-1) and 1 day post-infection community data. We identified 6 OTUs that
were important to the post-clindamycin model and either the baseline or
1 day post-infection models (\emph{Enterobacteriaceae, Ruminococcaceae,
Lactobacillus, Bacteroides, Porphyromonadaceae, Erysipelotrichaceae};
Table S14). Thus, a subset of bacterial OTUs were important for
determining \emph{C. difficile} colonization dynamics across multiple
timepoints.

@@

Applying our CDI model to 6 different sources of mice, allowed us to
identify bacterial taxa that were unique to different sources as well as
taxa that were universally impacted by clindamycin. We trained logistic
regression models with baseline (day -1), post-clindamycin treatment
(day 0), and 1-day post-infection fecal community data that could
predict whether mice cleared \emph{C. difficile} by 7 days
post-infection better than random chance. We identified
\emph{Bacteroides, Enterococcus, Enterobacteriaceae, Porphyromonadaceae}
(Fig. 6) as candidate bacteria within these communities that influenced
variation in \emph{C. difficile} colonization dynamics since these
bacteria were all important in the logistic regression models and varied
by source, were impacted by clindamycin treatment, or both. Overall, our
results demonstrated clindamycin was sufficient to render mice from
multiple sources susceptible to CDI and only a subset of the
inter-individual microbiota variation across mice from different sources
was needed to predict which mice could clear \emph{C. difficile}.

Other studies have used mice from multiple sources to identify bacteria
that either promote colonization resistance or increase susceptibility
to enteric infections (22, 23, 26--30). For example, against
\emph{Salmonella} infections, \emph{Enterobacteriaceae} and segmented
filamentous bacteria have emerged as protective (22, 27). We found
\emph{Enterobacteriaceae} increased in all sources of mice after
clindamycin treatment, facilitating \emph{C. difficile} colonization.
However, there was also variation in \emph{Enterobacteriaceae} relative
abundance levels between sources that was associated with the variation
in \emph{C. difficile} colonization dynamics across sources. Thus,
bacteria may have differential roles in determining susceptibility
depending on the type of bacterial infection.

Differences in CDI mouse model studies have been attributed to
intestinal microbiota variation across sources. For example, researchers
using the same clindamycin treatment and C57BL/6 mice had different
\emph{C. difficile} outcomes, one having sustained colonization (32),
while the other had transient colonization (18), despite both using
\emph{C. difficile} VPI 10643. Baseline differences in the microbiota
composition have been hypothesized to partially explain the differences
in colonization outcomes and overall susceptibility to \emph{C.
difficile} after treatment with the same antibiotic (13, 31). When we
treated mice from 6 different sources with clindamycin and challenged
them with \emph{C. difficile} 630, we found microbiota variation across
sources impacted colonization outcomes, but not susceptibility. A
previous study with \emph{C. difficile} identified an endogenous
protective \emph{C. difficile} strain LEM1 that bloomed after antibiotic
treatment in mice from Jackson or Charles River Laboratories, but not
Taconic that protected mice against the more toxigenic \emph{C.
difficile} VPI10463 (26). Given that we obtained mice from the same
vendors, we checked all mice for endogenous \emph{C. difficile} by
plating stool samples that were collected after clindamycin treatment.
However, we did not identify any endogenous \emph{C. difficile} strains
prior to challenge, suggesting there were no endogenous protective
strains in the mice we received and other bacteria mediated the
variation in \emph{C. difficile} colonization across sources. The
\emph{C. difficile} strain used could also be contributing to the
variation in \emph{C. difficile} outcomes seen across different research
groups. For example, a group found differential colonization outcomes
after clindamycin treatment, with \emph{C. difficile} 630 and M68
infections eventually becoming undetectable while strain BI-7 remained
detectable up to 70 days post-treatment (44). The bacterial
perturbations induced by clindamycin treatment have been well
characterized and our findings agree with previous CDI mouse model work
demonstrating \emph{Enterococcus} and \emph{Enterobacteriaceae} were
associated with \emph{C. difficile} susceptibility and
\emph{Porpyhromonadaceae}, \emph{Lachnospiraceae},
\emph{Ruminococcaceae}, and \emph{Turicibacter} were associated with
resistance (19, 21, 32, 33, 43--46). While we have demonstrated that
susceptibility is uniform across sources of mice after clindamycin
treatment, there could be different outcomes for either susceptibility
or clearance in the case of other antibiotic treatments.

@@

Using mice from multiple sources successfully increased the inter-animal
variation. One alternative approach that has been used in some CDI
studies is to associate mice with human microbiotas (47--52). However, a
major caveat to this method is the substantial loss of human microbiota
community members upon transfer to mice (53, 54). Additionally with the
exception of 2 recent studies (47, 48), most of these studies associated
mice with just 1 types of human microbiota either from a single donor or
a single pool from multiple donors (49--52). This approach does not aid
in the goal of modeling the interpersonal variation seen in humans to
understand how the microbiota influences susceptibility to CDIs and
adverse outcomes. Importantly, our study using mice from 6 different
sources increased the variation between groups of mice compared to using
1 source alone, to better reflect the inter-individual microbiota
variation observed in humans.

@@

Other microbiota and host factors that were outside the scope of our
current study may also contribute to the differences in \emph{C.
difficile} colonization dynamics between sources of mice. The microbiota
is composed of viruses, fungi, and parasites in addition to bacteria,
and these non-bacterial members can also vary across sources of mice
(55, 56). While our study focused solely on the bacterial portion,
viruses and fungi have also begun to be implicated in the context of
CDIs or FMT treatments for recurrent CDIs (35, 57--60). Beyond community
composition, the metabolic function of the microbiota also has a CDI
signature (20, 46, 61, 62) and can vary across mice from different
sources (63). For example, microbial metabolites, particularly secondary
bile acids and butyrate production, have been implicated as important
contributors to \emph{C. difficile} resistance (33, 44). Interestingly,
butyrate has previously been shown to vary across mouse vendors and
mediated resistance to \emph{Citrobacter rodentium} infection, a model
of enterohemorrhagic and enteropathogenic \emph{Escherichia coli}
infections (23). Evidence for immunological toning differences in IgA
and Th17 cells across mice from different vendors have also been
documented and (64, 65) could influence the host response to CDI (66,
67). The outcome after \emph{C. difficile} exposure depends on a
multitude of factors, including age, diet, and immunity; all of which
are also influenced by the microbiota.

@@

\textbf{(iii) \emph{C. difficile} infection model.} Mice were challenged
with 10\textsuperscript{3} spores of \emph{C. difficile} strain 630 via
oral gavage post-infection 1 day after clindamycin treatment as
described previously (21). Mice weights and stool samples were taken
daily through 9 days post-challenge (Fig. 1A). Collected stool was split
for \emph{C. difficile} quantification and 16S rRNA sequencing analysis.
For \emph{C. difficile} quantification, stool samples were transferred
to the anaerobic chamber, serially diluted in PBS, plated on
taurocholate-cycloserine-cefoxitin-fructose agar (TCCFA) plates, and
counted after 24 hours of incubation at 37°C under anaerobic conditions.
A sample from the day 0 timepoint (post-clindamycin and prior to
\emph{C. difficile} challenge) was also plated on TCCFA to ensure mice
were not already colonized with \emph{C. difficile} prior to infection.
There were 3 deaths recorded over the course of the experiment, 1
Taconic mouse died prior to \emph{C. difficile} challenge and 1 Jackson
and 1 Envigo mouse died between 1- and 3-days post-infection. Mice were
categorized as cleared when no \emph{C. difficile} was detected in the
first serial dilution (limit of detection: 100 CFU). Stool samples for
16S rRNA sequencing were snap frozen in liquid nitrogen and stored at
-80°C until DNA extraction.

\textbf{(iv) 16S rRNA sequencing.} DNA was extracted from -80°C stored
stool samples using the DNeasy Powersoil HTP 96 kit (Qiagen) and an
EpMotion 5075 automated pipetting system (Eppendorf). The V4 region was
amplified for 16S rRNA with the AccuPrime Pfx DNA polymerase (Thermo
Fisher Scientific) using custom barcoded primers, as previously
described (68). The ZymoBIOMICS microbial community DNA standards was
used as a mock community control (69) and water was used as a negative
control per 96-well extraction plate. The PCR amplicons were cleaned up
and normalized with the SequalPrep normalization plate kit (Thermo
Fisher Scientific). Amplicons were pooled and quantified with the KAPA
library quantification kit (KAPA biosystems), prior to sequencing using
the MiSeq system (Illumina).

>>>>

\hypertarget{the-initial-gut-microbiota-and-response-to-antibiotic-perturbation-influence-clostridioides-difficile-clearance-in-mice}{%
\section{\texorpdfstring{The initial gut microbiota and response to
antibiotic perturbation influence \emph{Clostridioides difficile}
clearance in
mice}{The initial gut microbiota and response to antibiotic perturbation influence Clostridioides difficile clearance in mice}}\label{the-initial-gut-microbiota-and-response-to-antibiotic-perturbation-influence-clostridioides-difficile-clearance-in-mice}}

@@

We sequenced the V4 region of the 16S rRNA gene from fecal samples
collected from these mice after they acclimated to the University of
Michigan animal housing environment. We first examined the alpha
diversity across the 6 sources of mice. There was a significant
difference in the richness (i.e.~number of observed operational
taxonomic units (OTUs)), but not Shannon diversity index across the
sources of mice (\emph{P}\textsubscript{FDR} = 0.03 and
\emph{P}\textsubscript{FDR} = 0.052, respectively; Fig. 1A-B and Tables
S1-2). Next, we compared the community structure of mice (Fig. 1C). The
source of mice and the interactions between the source and cage effects
explained most of the observed variation between fecal communities
(PERMANOVA combined R\textsuperscript{2} = 0.90, \emph{P} \textless{}
0.001; Fig. 1C and Table S3). Mice that are co-housed tend to have
similar gut microbiotas due to coprophagy (41). Since mice within the
same source were housed together, it was not surprising that the cage
effect also contributed to the observed community variation. There were
some differences between the 2 experiments we conducted, as the
experiment and cage effects significantly explained the observed
community variation for the Schloss and Young lab mouse colonies (Fig.
S1A-B and Table S4). However, most of the vendors also clustered by
experiment (Fig. S1C-D, F), suggesting there was some community
variation between the 2 experiments within each source, particularly for
Schloss, Young, and Envigo mice (Fig. S1G-H). After finding differences
at the community level, we next identified the bacteria that varied
between sources of mice. There were 268 OTUs with relative abundances
that were significantly different between the sources at baseline (Fig.
1D and Table S5). Though we saw differences between experiments at the
community level, there were no OTUs that were significantly different
between experiments within Schloss, Young, and Envigo mice at baseline
(all \emph{P} \textgreater{} 0.05). By using mice from six sources we
were able to increase the variation in the starting communities to
evaluate in a clindamycin-based CDI model.

\textbf{Clindamycin treatment renders all mice susceptible to \emph{C.
difficile} 630 colonization, but clearance time varies across sources.}
Clindamycin is frequently implicated with human CDIs (42) and was part
of the antibiotic treatment for the frequently cited 2008 CDI mouse
model (14). We have previously demonstrated mice are rendered
susceptible to \emph{C. difficile}, but clear the pathogen within 9
days, thus colonization is transient when treated with clindamycin alone
(21, 43). All mice were treated with 10 mg/kg clindamycin via
intraperitoneal injection and one day later challenged with
10\textsuperscript{3} \emph{C. difficile} 630 spores (Fig. 2A). The day
after infection, \emph{C. difficile} was detectable in all mice at a
similar level (median CFU range: 2.2e+07-1.3e+08;
\emph{P}\textsubscript{FDR} = 0.15), indicating clindamycin rendered all
mice susceptible regardless of source (Fig. 2B). However, between 3 and
7 days post-infection, we observed variation in \emph{C. difficile}
levels across sources of mice (all \emph{P}\textsubscript{FDR} \(\le\)
0.019; Fig. 2B and Table S6). This suggested the source of mice was
associated with \emph{C. difficile} clearance. While the colonization
dynamics were similar between the two experiments, the Schloss mice took
longer to clear \emph{C. difficile} in the first experiment compared to
the second and the Envigo mice took longer to clear \emph{C. difficile}
in the second experiment compared to the first (Fig. S2A-B). The change
in the mice's weight significantly varied across sources of mice with
the most weight loss occurring two days post-infection (Fig. 2C and
Table S7). There was also one Jackson and one Envigo mouse that died
between 1- and 3-days post-infection during the second experiment. Mice
obtained from Jackson, Taconic, and Envigo tended to lose more weight,
have higher \emph{C. difficile} CFU levels and take longer to clear the
infection compared to the other sources of mice (although there was
variation between experiments with Schloss and Envigo mice). This was
particularly evident 7 days post-infection (Fig. 2B-C, Fig. S2C-D), when
57\% of the mice were still colonized with \emph{C. difficile} (Fig.
S2E). By 9 days post-infection the majority of the mice from all sources
had cleared \emph{C. difficile} with the exception of 1 Taconic mouse
from the first experiment and 2 Envigo mice from the second experiment
(Fig. 2B). Thus, clindamycin rendered all mice susceptible to \emph{C.
difficile} 630 colonization, regardless of source, but there was
significant variation in disease phenotype across the sources of mice.

\textbf{Clindamycin treatment alters bacteria in all sources, but a
subset of bacterial differences across sources persists.} Given the
variation in fecal communities that we observed across breeding
colonies, we hypothesized that variation in \emph{C. difficile}
clearance would be explained by community variation across the 6 sources
of mice. As expected, clindamycin treatment decreased the richness and
Shannon diversity across all sources of mice (Fig. 3A-B). Interestingly,
significant differences in diversity metrics between sources emerged
after clindamycin treatment, with Charles River mice having higher
richness and Shannon diversity than most of the other sources
(\emph{P}\textsubscript{FDR} \textless{} 0.05; Fig 3A-B and Tables
S1-2). The clindamycin treatment decreased the variation in community
structures between sources of mice. The source of mice and the
interactions between source and cage effects explained almost all of the
observed variation between communities (combined R\textsuperscript{2} =
0.99, \emph{P} \textless{} 0.001; Fig. 3C and Table S3). However, there
were only 18 OTUs with relative abundances that significantly varied
between sources after clindamycin treatment (Fig. 3D and Table S8).
Next, we identified the bacteria that shifted after clindamycin
treatment, regardless of source by analyzing paired fecal samples from
mice that were collected at baseline and after clindamycin treatment. We
identified 153 OTUs that were altered after clindamycin treatment in
most mice (Fig. 3E and Table S9). When we compared the list of
significant clindamycin impacted bacteria with the bacteria that varied
between sources post-clindamycin, we found 4 OTUs that were shared
between the lists (\emph{Enterobacteriaceae (OTU 1)},
\emph{Lachnospiraceae} (OTU 130), \emph{Lactobacillus} (OTU 6),
\emph{Enterococcus} (OTU 23); Fig. 3D-E and Tables S8-9). Importantly,
some of the OTUs that varied between sources also shifted with
clindamycin treatment. For example, \emph{Proteus} increased after
clindamycin treatment (Fig. 3D), but only in Taconic mice.
\emph{Enterococcus} was primarily found in mice purchased from
commercial vendors and also increased in relative abundance after
clindamycin treatment (Fig. 3D). These findings demonstrate that
clindamycin had a consistent impact on the fecal bacterial communities
of mice from all sources and only a subset of the OTUs continued to vary
between sources.

@@

\textbf{Baseline, post-clindamycin, and post-infection community data
can predict mice that will clear \emph{C. difficile} by 7 days
post-infection.} After identifying taxa that varied between sources,
changed after clindamycin treatment, or both, we determined which taxa
were influencing the variation in \emph{C. difficile} colonization at
day 7 (Fig. 2B, Fig. S2C). We trained three L2-regularized logistic
regression models with either input bacterial community data from the 6
sources of mice at the baseline (day = -1), post-clindamycin (day = 0),
or post-infection (day = 1) timepoints of the experiment to predict
\emph{C. difficile} colonization status on day 7 (Fig. S3A-B). All
models were better at predicting \emph{C. difficile} colonization status
on day 7 than random chance (all \emph{P} \textless{} 0.001, Table S12).
The model based on the post-clindamycin (AUROC = 0.78) community OTU
data performed significantly better than the baseline (AUROC = 0.72) or
the post-infection (AUROC = 0.67) models (\emph{P}\textsubscript{FDR}
\textless{} 0.001 for pairwise comparisons; Fig. S3C and Table S13).
Thus, we were able to use bacterial relative abundance data from the
time of \emph{C. difficile} challenge to differentiate mice that had
cleared \emph{C. difficile} before day 7 from the mice still colonized
with \emph{C. difficile} at that timepoint. This result suggests that
the bacterial community's response to clindamycin treatment had the
greatest influence on subsequent \emph{C. difficile} colonization
dynamics.

To examine the bacteria that were driving each model's performance, we
selected the 20 OTUs that had the highest absolute feature weights in
each of the 3 models (Table S14). First, we looked at OTUs from the
model with the best performance, which was based on the post-clindamycin
treatment (day 0) bacterial community data. Out of the 10 highest ranked
OTUs, 7 OTUs were associated with \emph{C. difficile} colonization 7
days post-infection (\emph{Bacteroides}, \emph{Escherichia/Shigella}, 2
\emph{Lachnospiraceae}, \emph{Lactobacillus}, \emph{Porphyromonadaceae},
and \emph{Ruminococcaceae}), while 3 OTUs were associated with clearance
(\emph{Enterobacteriaceae}, \emph{Lachnospiraceae},
\emph{Porphyromonadaceae}; Fig. 5A). On day 0, the majority of these
OTUs were impacted by clindamycin and had relative abundances that were
close to the limit of detection (Fig. 5A). Next, we examined whether any
of the top 20 ranked OTUs from the post-clindamycin (day 0) model were
also important in the other 2 classification models based on baseline
(day -1) and 1 day post-infection community data. We identified 6 OTUs
that were important to the post-clindamycin model and either the
baseline or 1 day post-infection models (\emph{Enterobacteriaceae,
Ruminococcaceae, Lactobacillus, Bacteroides, Porphyromonadaceae,
Erysipelotrichaceae}; Table S14). Thus, a subset of bacterial OTUs were
important for determining \emph{C. difficile} colonization dynamics
across multiple timepoints.

@@

Applying our CDI model to 6 different sources of mice, allowed us to
identify bacterial taxa that were unique to different sources as well as
taxa that were universally impacted by clindamycin. We trained
L2-regularized logistic regression models with baseline (day -1),
post-clindamycin treatment (day 0), and 1-day post-infection fecal
community data that could predict whether mice cleared \emph{C.
difficile} by 7 days post-infection better than random chance. We
identified \emph{Bacteroides, Enterococcus, Enterobacteriaceae,
Porphyromonadaceae} (Fig. 6) as candidate bacteria within these
communities that influenced variation in \emph{C. difficile}
colonization dynamics since these bacteria were all important in the
logistic regression models and varied by source, were impacted by
clindamycin treatment, or both. Overall, our results demonstrated
clindamycin was sufficient to render mice from multiple sources
susceptible to CDI and only a subset of the inter-individual microbiota
variation across mice from different sources was needed to predict which
mice could clear \emph{C. difficile}.

Other studies have used mice from multiple sources to identify bacteria
that either promote colonization resistance or increase susceptibility
to enteric infections (22, 23, 26--30). For example, against
\emph{Salmonella} infections, \emph{Enterobacteriaceae} and segmented
filamentous bacteria have emerged as protective (22, 27). We found
\emph{Enterobacteriaceae} increased in all sources of mice after
clindamycin treatment, positively correlating with \emph{C. difficile}
colonization. However, there was also variation in
\emph{Enterobacteriaceae} relative abundance levels between sources that
was associated with the variation in \emph{C. difficile} colonization
dynamics across sources. Thus, bacteria may have differential roles in
determining susceptibility depending on the type of bacterial infection.

Differences in CDI mouse model studies have been attributed to
intestinal microbiota variation across sources. For example, researchers
using the same clindamycin treatment and C57BL/6 mice had different
\emph{C. difficile} outcomes, one having sustained colonization (32),
while the other had transient colonization (18), despite both using
\emph{C. difficile} VPI 10643. Baseline differences in the microbiota
composition have been hypothesized to partially explain the differences
in colonization outcomes and overall susceptibility to \emph{C.
difficile} after treatment with the same antibiotic (13, 31). When we
treated mice from 6 different sources with clindamycin and challenged
them with \emph{C. difficile} 630, we found microbiota variation across
sources impacted colonization outcomes, but not susceptibility. A
previous study with \emph{C. difficile} identified an endogenous
protective \emph{C. difficile} strain LEM1 that bloomed after antibiotic
treatment in mice from Jackson or Charles River Laboratories, but not
Taconic that protected mice against the more toxigenic \emph{C.
difficile} VPI10463 (26). Given that we obtained mice from the same
vendors, we checked all mice for endogenous \emph{C. difficile} by
plating stool samples that were collected after clindamycin treatment.
However, we did not identify any endogenous \emph{C. difficile} strains
prior to challenge, suggesting there were no endogenous protective
strains in the mice we received and other bacteria mediated the
variation in \emph{C. difficile} colonization across sources. The
\emph{C. difficile} strain used could also be contributing to the
variation in \emph{C. difficile} outcomes seen across different research
groups. For example, a group found differential colonization outcomes
after clindamycin treatment, with \emph{C. difficile} 630 and M68
infections eventually becoming undetectable while strain BI-7 remained
detectable up to 70 days post-treatment (44). One study limitation is
that we only used female mice. Sex has been shown to influence
microbiota variation in mice (45), so we used female mice to reduce this
confounding variable and also match the sex used in previous CDI studies
that administered clindamycin to mice (32, 32, 44, 46). The bacterial
perturbations induced by clindamycin treatment have been well
characterized and our findings agree with previous CDI mouse model work
demonstrating \emph{Enterococcus} and \emph{Enterobacteriaceae} were
associated with \emph{C. difficile} susceptibility and
\emph{Porpyhromonadaceae}, \emph{Lachnospiraceae},
\emph{Ruminococcaceae}, and \emph{Turicibacter} were associated with
resistance (19, 21, 32, 33, 43, 44, 46, 47). While we have demonstrated
that susceptibility is uniform across sources of mice after clindamycin
treatment, there could be different outcomes for either susceptibility
or clearance in the case of other antibiotic treatments.

@@

Using mice from multiple sources successfully increased the inter-animal
variation. One alternative approach that has been used in some CDI
studies is to associate mice with human microbiotas (48--53). However, a
major caveat to this method is the substantial loss of human microbiota
community members upon transfer to mice (54, 55). Additionally, with the
exception of 2 recent studies (48, 49), most of these studies associated
mice with just 1 type of human microbiota either from a single donor or
a single pool from multiple donors (50--53). This approach does not aid
in the goal of modeling the interpersonal variation seen in humans to
understand how the microbiota influences susceptibility to CDIs and
adverse outcomes. Importantly, our study using mice from 6 different
sources increased the variation between groups of mice compared to using
1 source alone, to better reflect the inter-individual microbiota
variation observed in humans.

@@

Other microbiota and host factors that were outside the scope of our
current study may also contribute to the differences in \emph{C.
difficile} colonization dynamics between sources of mice. The microbiota
is composed of viruses, fungi, and parasites in addition to bacteria,
and these non-bacterial members can also vary across sources of mice
(56, 57). While our study focused solely on the bacterial portion,
viruses and fungi have also begun to be implicated in the context of
CDIs or FMT treatments for recurrent CDIs (35, 58--61). Beyond community
composition, the metabolic function of the microbiota also has a CDI
signature (20, 47, 62, 63) and can vary across mice from different
sources (64). For example, microbial metabolites, particularly secondary
bile acids and butyrate production, have been implicated as important
contributors to \emph{C. difficile} resistance (33, 44). Interestingly,
butyrate has previously been shown to vary across mouse vendors and
mediated resistance to \emph{Citrobacter rodentium} infection, a model
of enterohemorrhagic and enteropathogenic \emph{Escherichia coli}
infections (23). Evidence for immunological toning differences in IgA
and Th17 cells across mice from different vendors have also been
documented (65, 66) and could influence the host response to CDI (67,
68), particularly relevant for \emph{C. difficile} strains that induce
more severe disease than \emph{C. difficile} 630. The outcome after
\emph{C. difficile} exposure depends on a multitude of factors,
including genetics, age, diet, and immunity; all of which also influence
the microbiota.

@@

\textbf{(iii) \emph{C. difficile} infection model.} Mice were challenged
with 10\textsuperscript{3} spores of \emph{C. difficile} strain 630 via
oral gavage post-infection 1 day after clindamycin treatment as
described previously (21). Mice weights and stool samples were taken
daily through 9 days post-infection (Fig. 1A). Collected stool was split
for \emph{C. difficile} quantification and 16S rRNA sequencing analysis.
For \emph{C. difficile} quantification, stool samples were transferred
to the anaerobic chamber, serially diluted in PBS, plated on
taurocholate-cycloserine-cefoxitin-fructose agar (TCCFA) plates, and
counted after 24 hours of incubation at 37°C under anaerobic conditions.
A sample from the day 0 timepoint (post-clindamycin and prior to
\emph{C. difficile} challenge) was also plated on TCCFA to ensure mice
were not already colonized with \emph{C. difficile} prior to infection.
There were 3 deaths recorded over the course of the experiment, 1
Taconic mouse died prior to \emph{C. difficile} challenge and 1 Jackson
and 1 Envigo mouse died between 1- and 3-days post-infection. Mice were
categorized as cleared when no \emph{C. difficile} was detected in the
first serial dilution (limit of detection: 100 CFU). Stool samples for
16S rRNA sequencing were snap frozen in liquid nitrogen and stored at
-80°C until DNA extraction.

\textbf{(iv) 16S rRNA sequencing.} DNA was extracted from -80°C stored
stool samples using the DNeasy Powersoil HTP 96 kit (Qiagen) and an
EpMotion 5075 automated pipetting system (Eppendorf). The V4 region was
amplified for 16S rRNA with the AccuPrime Pfx DNA polymerase (Thermo
Fisher Scientific) using custom barcoded primers, as previously
described (69). The ZymoBIOMICS microbial community DNA standards was
used as a mock community control (70) and water was used as a negative
control per 96-well extraction plate. The PCR amplicons were cleaned up
and normalized with the SequalPrep normalization plate kit (Thermo
Fisher Scientific). Amplicons were pooled and quantified with the KAPA
library quantification kit (KAPA biosystems), prior to sequencing using
the MiSeq system (Illumina).
